---
title: "What is the Best Way to Promote Zero Calories Beverage"
format: html
editor: visual
---

## 3. Exploratory Data Analysis

```{r, echo=FALSE, results='hide'}
library(here)
library(dplyr)
data <- read_csv(here("rawdata/june1data.csv"))
data <- filter(data, Total > 0) 
data <- data[,-c(7,8,9)]
```

Specific to the statistical questions listed in introduction, the following exploratory data analysis procedures are suggested. Firstly, since how consumption of zero calories beverage is affected by site is of interest of study, a "sphagetti plot", which generally refers to a plot showing the changes of multiple flows over time, is suggested to visualize how zero calories consumption changes in different locations, and an example plot is as following:

```{r figs, echo=FALSE, fig.width=7,fig.height=4,fig.cap="\\label{fig:figs}plotting example"}
library(ggplot2)
ggplot(data=data, aes(x=Count, y=ZeroCal, color=Site))+
  geom_point()+
  geom_line()+
  labs(x="Days of Experiment", y="Bottles of Consumption",
       title="The Sphagetti Plot of Zero Calories Beverage Consumption Against Time")+
  theme(plot.title=element_text(hjust=0.5))
```