---
title: "What is the best way to promote zero calories beverages?"
format: 
  pdf:
    fontsize: 11pt
editor: visual
bibliography: ../reference/reference.bib
author: Ivy Liu, Zefan Liu, Tom Tang
---

## Introduction

Zero calorie beverages are healthier compared to sugared beverages because they contain less sugar and significantly fewer calories. Over the past decades, hospitals have been striving to promote zero-calorie beverages for a healthier lifestyle. This study employs five interventions to help increase such promotion. These interventions include a 10% price discount, a price discount combined with messaging that explains the reason for the discount, messaging displaying the caloric content in sugared beverages, messaging indicating the amount of exercise needed to burn off the calories in a sugared beverage, and messaging containing both caloric content and the amount of physical activity needed.

The study aims to investigate whether these interventions help encourage people to choose zero calorie beverages over sugared beverages and if such effects varies across hospitals.

## Data description and summaries

This is an experimental design where data are collected over 30 weeks, from October 27 to May 23, with a follow-up period of 14 days, gathered from two urban hospitals and one suburban hospital. During this time period, the sales of different types of drinks are recorded (see Table \ref{tab1}).

```{=tex}
\begin{table}[htbp]
    \centering
    \caption{Summary of data}
    \label{tab1}
    \begin{tabular}{ c|c|c }
        \hline
        Data Name & Data Type & Data Description \\
        \hline
        Count        & Categorical & The day from the beginning of study
        DofW         & Categorical & The day of the week, 7 levels \\
        Site         & Categorical & The location of the hospital, 3 levels \\
        Intervention & Categorical & The intervention applied, 9 levels \\
        ZeroCal      & Integer     & The number of zero calorie drinks sold \\
        Sugary       & Integer     & The number of sugary drinks sold \\
        Juice100     & Integer     & The number of Juice100 sold \\
        Ojuice       & Integer     & The number of Ojuice sold \\
        Sports       & Integer     & The number of sports drinks sold \\
        Total        & Continuous  & The sum of all drinks sold \\
        \hline
    \end{tabular}
\end{table}
```
There are missing entries in this dataset, with a significant portion of the missing data attributed to the 'Juice100,' 'Ojuice,' and 'Sports' columns. Since neither these measurements themselves nor their links with the consumption of ZeroCal/Sugary beverages are the main focus of the study, the missing values in these columns should not be given specific attention. Additionally, there are seven entirely missing consecutive entries at HF hospital during the follow-up period and two isolated cases of missing entries. Such occurrences could be due to public holidays or temporary closures for renovation. In general, such missing data can be classified as missing at random.

## Exploratory Data Analysis

```{r, include=FALSE}
library(here)
library(readr)
library(dplyr)
data <- read_csv(here("rawdata/june1data.csv"))
data <- filter(data, Total > 0) 
data <- data[,-c(7,8,9)]
```

Firstly, to address the influence of location on zero-calorie beverage consumption, which is of primary interest to the study, a spaghetti plot is recommended. A spaghetti plot is typically used to demonstrate the change of multiple flows over time and how they vary across the three sites. Based on Figure 1, it is evident that the time-series data for zero-calorie beverage consumption at the site "chop" are significantly higher than those at the other two sites, suggesting substantial variation in consumption distributions under different interventions. Additionally, the spaghetti plot reveals that except for the initial period at the site "HF," there is no apparent trend at all three sites, as the long-term averages seem to be stable. However, strong weekly variations are evident, implying that it is reasonable to include the day of the week (DofW) and exclude the time (count) in the model.

```{r figs, echo=FALSE, fig.width=7,fig.height=4,fig.cap="\\label{fig:figs}Examples of Recommended EDA Plots"}
library(ggplot2)
library(cowplot)
ggplot(data=data, aes(x=Count, y=ZeroCal, color=Site))+
  geom_point()+
  geom_line()+
  labs(x="Days of Experiment", y="ZeroCal (bottle)",
       title="The Spaghetti Plot")+
  theme(plot.title=element_text(hjust=0.5),
        legend.position="bottom")

```

## Formal analysis

It is recommended to implement a Generalized Linear Mixed Model (GLMM) to address the problems. A GLMM, an extension of the Linear Mixed Model (LMM) [@winter2013linear], can incorporate both fixed and random effects, and also accommodate the response variable being a count variable through a link function [@dobson2018introduction]. Specifically, while the fixed effects assess the association of covariates with the response variable across the overall population---serving as a baseline---the random effects account for variations in some associations across different entities or individuals. Link function can guarantee that the range of model outputs matches with the range of values that can be observed, which is the advantage of GLMM over LMM.

In this study, the primary focus is to analyze the effects of the interventions, and therefore, it is suggested to treat Intervention as a fixed effect. Additionally, DofW should also be considered as fixed effects because they apply to all three Sites. Random effect should be Site, which characterizes the objects being observed over time. Note that although Total is also a reasonable predictor for both variables of interest, it is more appropriate to place it as a offset term in model, whose effect to model is "structural" with a fixed coefficient of one. Given that the consumption is non-negative integer in data, a natural choice of link function is to apply logarithm transformation on ZeroCal, Sugary and Total, which guarantees that model outputs are positive.

Hence, the proposed models are

```{=tex}
\begin{align*}
\log(\text{ZeroCal}_{ij}) &= (\beta_0+b_i) + \beta_1\text{Intervention} + \beta_2\text{DofW} + \text{offset}(\log(Total))\\
\log(\text{Sugary}_{ij}) &= (\beta_0+b_i) + \beta_1\text{Intervention} + \beta_2\text{DofW} + \text{offset}(\log(Total))
\end{align*}
```
where $b_i \sim N(0,\sigma^2)$ represents the random effect from Site. Note that response variables, which are counts, can be assumed to follow either Poisson or Negative Binomial distribution under log link function. The former is more common but also more restrictive.

```{r, include=FALSE}
library(lme4)
library(glmmTMB)
data$Site <- as.factor(data$Site)
data$Intervention <- as.factor(data$Intervention)
data$Intervention <- relevel(data$Intervention, 
                             ref = "preint")
data$DofW <- as.factor(data$DofW)

zerocal_lme = glmer(ZeroCal ~ (1|Site) + Intervention 
                    + DofW + offset(log(Total)), data=data, family=poisson)
zerocal_lme_nb = glmmTMB(ZeroCal ~ (1|Site) + offset(log(Total)) + Intervention 
                        + DofW, data=data, family=nbinom1)
AIC(zerocal_lme)
AIC(zerocal_lme_nb)

sugary_lme = glmer(Sugary ~ (1|Site) + offset(log(Total)) + Intervention 
                    + DofW, data=data, family=poisson)
sugary_lme_nb = glmmTMB(Sugary ~ (1|Site) + offset(log(Total)) + Intervention 
                        + DofW, data=data, family=nbinom1)
AIC(sugary_lme)
AIC(sugary_lme_nb)
```

Based on our analysis, Generalized Mixed Effect Models assuming response variables follow negative binomial distribution have the best performance in terms of Akaike information criterion (AIC), which is one of the most common criterions for model selection. For the consumption of Zero-Cal beverage, our model suggests that only the effect of the intervention "discount+messaging" are statistically significant; for the consumption of Sugary beverage, our model also suggests that the effect of the intervention "discount+messaging" is significant. Besides, the effects from "DofW=6" and "DofW=7" are also significant, implying that on average the consumption of sugary beverage may differ between weekdays and weekend.

## Conclusions and Limitation



## References

::: {#refs}
:::

## Statistical Appendix

### Mathematical Formulation of Models

Here is a formulation of the Generalized Linear Mixed Model within the context of this research. As suggested earlier, random effects can be placed on both the intervention's slope and intercept. Taking $y = ZeroCal, x_1 = Total, x_2 = Intervention, x_3 = DofW$, the model will take the format of

```{=tex}
\begin{equation}
\label{eq1}
E(y_{ij}) = exp((\beta_0 + b_{0i}) + (\beta_1 + b_{1i} + b_1 x_1)\; x_2 + \beta_2 \; x_3 + \beta_3 \; x_1)
\end{equation}
```
where $b_{0i} \sim N(0,\sigma_1^2)$, $b_{1i} \sim N(0,\sigma_2^2)$ and $b_{1} \sim N(0,\sigma_3^2)$. Here, $i \in \{1,2,3\}$ is used to index the three sites, and $j\in \{1,2, \dots, 221\}$ is used to index the date. $\sigma_1$, $\sigma_2$ and $\sigma_3$ can be estimated using data, and $zero\_calorie_{ij}$ is assumed to have poisson distribution where its expectation should be equal to its variance. When such an assumption is violated, usually quasi-maximum likelihood estimation is used. This method estimates the regression parameters without the need to specify a distribution for the outcome. When such an assumption is violated, usually quasi-maximum likelihood estimation is used. This method estimates the regression parameters without the need to specify a distribution for the outcome.
