---
title: "What is the best way to promote zero calories beverages?"
format: 
  pdf:
    fontsize: 11pt
editor: visual
author: Ivy Liu, Zefan Liu, Tom Tang
---

## Introduction

Zero calorie beverages are healthier compared to sugared beverages because they contain less sugar and significantly fewer calories. Over the past decades, hospitals have been striving to promote zero-calorie beverages for a healthier lifestyle. This study employs five interventions to help increase such promotion. These interventions include a 10% price discount, a price discount combined with messaging that explains the reason for the discount, messaging displaying the caloric content in sugared beverages, messaging indicating the amount of exercise needed to burn off the calories in a sugared beverage, and messaging containing both caloric content and the amount of physical activity needed.

The study aims to investigate whether these interventions help encourage people to choose zero calorie beverages over sugared beverages and if such effects are consistent across hospitals. The primary statistical question in this context is: "Is there any association between interventions and the increase in zero calorie beverage consumption or decrease in sugary beverage consumption?" The secondary statistical question is:"Does the effect of intervention differs by site?"

## Data description and summaries

This is an experimental design where data are collected over 30 weeks, from October 27 to May 23, with a follow-up period of 14 days, gathered from two urban hospitals and one suburban hospital. During this time period, the sales of different types of drinks are recorded.

| Data Name    | Data Type   | Data Description                       |
|--------------|-------------|----------------------------------------|
| Count        | Integer     | The number of days                     |
| DofW         | Categorical | The day of the week, 7 levels          |
| Site         | Categorical | The location of the hospital, 3 levels |
| Intervention | Categorical | The intervention applied, 9 levels     |
| ZeroCal      | Integer     | The number of zero calorie drink sold  |
| Sugary       | Integer     | The number of sugary drink sold        |
| Juice100     | Integer     | The number of Juice100 sold            |
| Ojuice       | Integer     | The number of Ojuice sold              |
| Sports       | Integer     | The number of sports drink sold        |
| Total        | Continuous  | The sum of all drink sold              |

There are missing entries in this dataset, with a significant portion of the missing data attributed to the 'Juice100,' 'Ojuice,' and 'Sports' columns. These columns will not be included in the model because of their irrelevancy, so the missing entries in these columns will not affect the statistical analysis. Additionally, there are seven entirely missing consecutive entries at HF hospital during the follow-up period. This is possible due to the hospital's cafeteria and convenient shop close down for renovation. Apart from these, there are two isolated cases of missing entries. These can be reasonably attributed to temporary closures for public holidays.

## Exploratory Data Analysis

```{r, include=FALSE}
library(here)
library(readr)
library(dplyr)
data <- read_csv(here("rawdata/june1data.csv"))
data <- filter(data, Total > 0) 
data <- data[,-c(7,8,9)]
```

Specific to the statistical questions listed in the introduction, the following exploratory data analysis procedures are suggested. Firstly, in order to address the influence of location on zero-calorie beverage consumption, which is of primary interest to the study, a spaghetti plot is recommended. A spaghetti plot is typically used to demonstrate the change of multiple flows over time and how it varies across the three sites. Meanwhile, it is recommended to use boxplots to approach the primary statistical question concerning the impact of different interventions on zero-calorie and sugary beverage consumption. Boxplots enable a preliminary check of consumption distributions under various interventions. Illustrated below are examples of a spaghetti plot and a boxplot (note that for demonstration purposes, only three interventions are included in the example boxplot).

```{r figs, echo=FALSE, fig.width=7,fig.height=4,fig.cap="\\label{fig:figs}Examples of Recommended EDA Plots"}
library(ggplot2)
library(cowplot)
p1=ggplot(data=data, aes(x=Count, y=ZeroCal, color=Site))+
  geom_point()+
  geom_line()+
  labs(x="Days of Experiment", y="ZeroCal (bottle)",
       title="The Spaghetti Plot")+
  theme(plot.title=element_text(hjust=0.5),
        legend.position="bottom")

example_data <- data |> filter(Intervention %in% 
                                 c("cal", "dismes", "preint"))

p2=ggplot(data=example_data, aes(x=Intervention, y=ZeroCal))+
  geom_boxplot(outlier.colour="red", outlier.shape=8,
                outlier.size=4)+
  labs(x="Interventions", y="ZeroCal (bottle)",
       title="The Box-plot")+
  theme(plot.title=element_text(hjust=0.5),
        legend.position="bottom")

plot_grid(p1,p2, ncol=2)
```

Based on the two plots above, it's clear that the time-series data for zero-calorie beverage consumption at site "chop" significantly diverges from the patterns observed at the other two sites, and that the distributions of consumption may vary substantially under different interventions. Additionally, there are two observations. Firstly, the spaghetti plot reveals that except for the beginning period at site "HF", there is no apparent trend at all three sites as the long-term averages seem to be stable. However, strong weekly variations are presented, implying that it is reasonable to include the day of the week (DofW) and to exclude the time (count) in the model. Secondly, the box plot indicates the presence of many outliers during which the caloric messaging was displayed. However, the clustering nature of these outliers tells valuable information. Therefore, it is advised that outliers be retained in the analysis to reflect the true variability inherent in the intervention's effects.

## Formal analysis

This section outlines general strategies for quantifying and comparing the impacts of five different interventions on the consumption of zero-calorie and sugary beverages across three sites.

It is recommended to implement a Linear Mixed Model (LMM) to approach the problems. An LMM is a statistical model that can incorporate both fixed effects and random effects. In particular, the fixed effects are typically the primary variables of interest that are representative of the entire population, while the random effects are parameters that vary across different hierarchical levels of the data, measuring variability between individuals.

In this study, the primary focus is to analyze the effects of the interventions, and therefore it is suggested to treat the intervention as a fixed effect. Additionally, days of the week should also be considered as a fixed effect because the weekly fluctuation applies to all three sites. To allow for a more comprehensive understanding of the intervention effects, it is advised to include sites and total daily beverage consumption as random effects in the model. These covariates control for variability that might influence the patterns of zero-calorie/sugary beverage consumption. For instance, site differences (which are the secondary statistical question) may impact purchasing behavior due to demographic variations between urban and suburban areas. Similarly, accounting for total daily beverage consumption helps isolate the effect of human traffic on zero-calorie/sugary beverage choices. This comprehensive approach ensures that the analysis accurately reflects the interventions' impact, while controlling for external factors.

Hence, the proposed models are $ZeroCal \sim Intervention + DofW + Site + Total$ and $Sugary \sim Intervention + DofW + Site + Total$, where $Intervention$ and $DofW$ are treated as fixed effects, and the others as random effects. Specifically, the coefficients for $Intervention$ measures the effects of corresponding interventions to the response variables, while controlling the covariates in the model. Interpreting the coefficients for $Intervention$ is crucial for answering the statistical questions.

## Conclusions

We recommend using a linear mixed-effects model and fitting two models. The first model uses ZeroCal as a response variable. Our fixed-effect explanatory variables are intervention and DofW, and our random-effect explanatory variables are Site and Total. This model will be useful to assess any association between our explanatory variables and ZeroCal. The second model uses Sugary as a response variable, with the same setup as the previous model. The second model will be ideal for demonstrating any association between the explanatory variables and Sugary.

## References

## Statistical Appendix

$$
zero\_calorie_{ij} = (\beta_0 + b_{0i}) + (\beta_1 + b_{1i}) \;  site + \beta_2 \; Intervention + \beta_3 \; DofW + \beta_4 \; Total
$$ $$
sugary_{ij} = (\beta_0 + b_{0i}) + (\beta_1 + b_{1i}) \;  site + \beta_2 \; Intervention + \beta_3 \; DofW + \beta_4 \; Total
$$
