---
title: "What is the best way to promote zero calories beverages?"
format: 
  pdf:
    fontsize: 11pt
editor: visual
bibliography: ../reference/reference.bib
author: Ivy Liu, Zefan Liu, Tom Tang
---

## Introduction

Zero calorie beverages are healthier compared to sugared beverages because they contain less sugar and significantly fewer calories. Over the past decades, hospitals have been striving to promote zero-calorie beverages for a healthier lifestyle. This study employs five interventions to help increase such promotion. These interventions include a 10% price discount, a price discount combined with messaging that explains the reason for the discount, messaging displaying the caloric content in sugared beverages, messaging indicating the amount of exercise needed to burn off the calories in a sugared beverage, and messaging containing both caloric content and the amount of physical activity needed.

The study aims to investigate whether these interventions help encourage people to choose zero calorie beverages over sugared beverages and if such effects varies across hospitals.

## Data description and summaries

This is an experimental design where data are collected over 30 weeks, from October 27 to May 23, with a follow-up period of 14 days, gathered from two urban hospitals and one suburban hospital. During this time period, the sales of different types of drinks are recorded (see Table \ref{tab1}).

```{=tex}
\begin{table}[htbp]
    \centering
    \caption{Summary of data}
    \label{tab1}
    \begin{tabular}{ c|c|c }
        \hline
        Data Name & Data Type & Data Description \\
        \hline
        DofW         & Categorical & The day of the week, 7 levels \\
        Site         & Categorical & The location of the hospital, 3 levels \\
        Intervention & Categorical & The intervention applied, 9 levels \\
        ZeroCal      & Integer     & The number of zero calorie drinks sold \\
        Sugary       & Integer     & The number of sugary drinks sold \\
        Juice100     & Integer     & The number of Juice100 sold \\
        Ojuice       & Integer     & The number of Ojuice sold \\
        Sports       & Integer     & The number of sports drinks sold \\
        Total        & Continuous  & The sum of all drinks sold \\
        \hline
    \end{tabular}
\end{table}
```

There are missing entries in this dataset, with a significant portion of the missing data attributed to the 'Juice100,' 'Ojuice,' and 'Sports' columns. Since neither these measurements themselves nor their links with the consumption of ZeroCal/Sugary beverages are the main focus of the study, the missing values in these columns should not be given specific attention. Additionally, there are seven entirely missing consecutive entries at HF hospital during the follow-up period and two isolated cases of missing entries. Such occurrences could be due to public holidays or temporary closures for renovation. In general, such missing data can be classified as missing at random.

## Exploratory Data Analysis

```{r, include=FALSE}
library(here)
library(readr)
library(dplyr)
data <- read_csv(here("rawdata/june1data.csv"))
data <- filter(data, Total > 0) 
data <- data[,-c(7,8,9)]
```

Firstly, to address the influence of location on zero-calorie beverage consumption, which is of primary interest to the study, a spaghetti plot is recommended. A spaghetti plot is typically used to demonstrate the change of multiple flows over time and how they vary across the three sites. Based on Figure 1, it is evident that the time-series data for zero-calorie beverage consumption at the site "chop" are significantly higher than those at the other two sites, suggesting substantial variation in consumption distributions under different interventions. Additionally, the spaghetti plot reveals that except for the initial period at the site "HF," there is no apparent trend at all three sites, as the long-term averages seem to be stable. However, strong weekly variations are evident, implying that it is reasonable to include the day of the week (DofW) and exclude the time (count) in the model.

```{r figs, echo=FALSE, fig.width=7,fig.height=4,fig.cap="\\label{fig:figs}Examples of Recommended EDA Plots"}
library(ggplot2)
library(cowplot)
ggplot(data=data, aes(x=Count, y=ZeroCal, color=Site))+
  geom_point()+
  geom_line()+
  labs(x="Days of Experiment", y="ZeroCal (bottle)",
       title="The Spaghetti Plot")+
  theme(plot.title=element_text(hjust=0.5),
        legend.position="bottom")

```

## Formal analysis

It is recommended to implement a Generalized Linear Mixed Model (GLMM) to address the problems. A GLMM, an extension of the Linear Mixed Model (LMM) [@winter2013linear], can incorporate both fixed and random effects, and also accommodate the response variable being a count variable through a link function [@dobson2018introduction]. Specifically, while the fixed effects assess the association of covariates with the response variable across the overall population---serving as a baseline---the random effects account for variations in some associations across different entities or individuals.

In this study, the primary focus is to analyze the effects of the interventions, and therefore, it is suggested to treat the intervention as a fixed effect. Additionally, days of the week should also be considered as fixed effects because they apply to all three sites. To allow for a more comprehensive understanding of the intervention effects, it is advised to include sites and total as two random effect in the model. These covariates control for variability that might influence the patterns of zero-calorie/sugary beverage consumption. For instance, site differences (which are the secondary statistical question) may impact purchasing behavior due to demographic variations between urban and suburban areas. Similarly, accounting for total daily beverage consumption helps isolate the effect of human traffic on zero-calorie/sugary beverage choices. Lastly, given that the consumption of ZeroCal beverage is count (non-negative integer), it is natural to assume that it follows a Poisson distribution. Subsequently, using an exponential link function to guarantee a positive output from model is suggested.

Hence, the proposed models are \begin{align*}
ZeroCal & \sim Intervention(Site) + DofW + Total\\
Sugary & \sim Intervention(Site) + DofW + Total.
\end{align*} Fixed effects include Intervention, DofW, and Total, while the random effect measures how the effect of Intervention and baseline consumption varies across sites. Note that these models are simplified versions, especially with the omission of the link function. Please refer to the appendix for the formulated models.

However, there are several limitations to the models mentioned above. Firstly, GLLMs do not accommodate autocorrelation in time series data, which refers to the dependence between consecutive observations. Moreover, as mentioned, the model proposed implicitly assumes that the response variables follow Poisson distributions [@coxe2009analysis], where the expectation and variability of response variables should equate, but this assumption can be easily violated in practice.

## Conclusions

We recommend using GLMMs to fit the relationship between covariates and response variables. Based on different research interests, response variable should be either ZeroCal beverage consumption or sugary beverage consumption. The fixed-effect explanatory variables suggested are intervention, Total and DofW, and the random-effect explanatory variable is Site.

## References

::: {#refs}
:::

## Statistical Appendix

### Mathematical Formulation of Models

There are two types of generalized linear mixed models fitted in this report, adapted for different structures of data. For illustration purpose, only the two models for zerocal beverage sales are shown. The first model (Equation \ref{eq1}) is GLMM with Poisson distribution, which is suitable for count data where the counts are grouped by random effects. In this case, the random effect is the site.

```{=tex}
\begin{equation}
\label{eq1}
\log(\text{Zerocal}_{ijk}) = \beta_0 + \beta_1 \cdot \text{Intervention}_i + \beta_2 \cdot \text{DoW} + \log(\text{Total}_{ijk}) + u_{j}
\end{equation}
```

where "Zerocal" is the count for the daily sales of zerocal beverages, "DoW" encodes day of the week, and "Total" is the daily sales of all beverages. In addition, $\beta_0$ is the intercept, $\beta_1$ is the coefficient for intervention, $\beta_2$ is the coefficient for "DoW", and $u_j$ is the random effect for the j-th Site. Notice that $log(\text{Total})$ is regarded as an offset because we want its coefficient to be fixed at 1.

The second model (Equation \ref{eq2}) is a GLMM with negative binomial distribution. Similar to the first model, it is also used for count data, but is more suitable when there is overdispersion (the variance is greater than the mean).
```{=tex}
\begin{equation}
\label{eq2}
\log(\text{Zerocal}_{ijk}) = \beta_0 + \beta_1 \cdot \text{Intervention}_i + \beta_2 \cdot \text{DoW} + \log(\text{Total}_{ijk}) + u_{j}
\end{equation}
```

### Model Selection
It is suggested in the formal analysis section that model should be selected based on AIC criterion. Table \ref{tab2} shows the AIC comparison.
```{=tex}
\begin{table}[htbp]
    \centering
    \caption{Model selection based on AIC criterion}
    \label{tab2}
    \begin{tabular}{ c|c|c }
        \hline
        GLMM Model & Zerocal beverage sales & Sugary beverage sales \\
        \hline
        Poisson            & 14874.39 & 12096.9 \\
        Negative Binomial  & 6765.097 & 6529.67 \\
        \hline
    \end{tabular}
\end{table}
```
THe GLMMs using negative binomial distribution are suggested because they give lower AIC scores.


### Model Diagnostic

```{r,include=FALSE}
library(lme4)
library(glmmTMB)
library(DHARMa)
library(here)
library(readr)
library(dplyr)
data <- read_csv(here("rawdata/june1data.csv"))
data$Site <- as.factor(data$Site)
data$Intervention <- as.factor(data$Intervention)
data$DofW <- as.factor(data$DofW)
```

```{r, echo=FALSE}
zerocal_lme_nb = glmmTMB(ZeroCal ~ (1|Site) + offset(log(Total)) + Intervention + DofW, data=data, family=nbinom1)
output <- simulateResiduals(fittedModel = zerocal_lme_nb, plot = F)
plotQQunif(output) 
plotResiduals(output)
```
These two plots correspond to the zero calories model.    

```{r, echo=FALSE}
sugary_lme_nb = glmmTMB(Sugary ~ (1|Site) + offset(log(Total)) + Intervention 
                        + DofW, data=data, family=nbinom1)
output1 <- simulateResiduals(fittedModel = sugary_lme_nb, plot = F)
plotQQunif(output1) 
plotResiduals(output1)
```
