---
title: "What is the best way to promote zero calories beverages?"
format: 
  pdf:
    fontsize: 11pt
editor: visual
author: Ivy Liu, Zefan Liu, Tom Tang
---

## Introduction



## Data description and summaries



## Exploratory Data Analysis

```{r, include=FALSE}
library(here)
library(readr)
library(dplyr)
data <- read_csv(here("rawdata/june1data.csv"))
data <- filter(data, Total > 0) 
data <- data[,-c(7,8,9)]
```

Specific to the statistical questions listed in introduction, the following exploratory data analysis procedures are suggested. Firstly, since how consumption of zero calories beverage is affected by site is of interest of study, a "spaghetti plot", which generally refers to a plot showing the changes of multiple flows over time, is suggested to visualize how zero calories consumption changes in different locations; meanwhile, for the primary statistical question, which is how different interventions would affect the consumption of zero calories beverage, it is suggested to use box-plot to check the distributions of the consumption under different interventions as a preview. Examples of a spaghetti plot and a box-plot is given below, and note that for convenience only three interventions are used in the example box-plot. 

```{r, echo=FALSE}
library(ggplot2)
library(cowplot)
p1=ggplot(data=data, aes(x=Count, y=ZeroCal, color=Site))+
  geom_point()+
  geom_line()+
  labs(x="Days of Experiment", y="Bottles of Consumption",
       title="Spaghetti Plot")+
  theme(plot.title=element_text(hjust=0.5))

example_data <- data |> filter(Intervention %in% 
                                 c("cal", "dismes", "preint"))

p2=ggplot(data=example_data, aes(x=Intervention, y=ZeroCal))+
  geom_boxplot(outlier.colour="red", outlier.shape=8,
                outlier.size=4)+
  labs(x="Interventions", y="Bottles of Consumption",
       title="Box Plot")+
  theme(plot.title=element_text(hjust=0.5),
        legend.position="top")

plot_grid(p1,p2, ncol=2)
```
From the two plots, we can clearly see that the time-series data of zero calories beverage consumption in Site=chop is pretty different from the consumption in the other two sites, and the distributions of consumption may vary alot under different interventions. Besides, there are two things to note. First of all, according to the spaghetti plot, except for the beginning period for Site=HF, there is no apparent trend in all three sites, but there is strong periodic variations, which implies that it is reasonable including DofW into covariates and removing count, since long term averages seems to be stable. Secondly, based on the box-plot, there are many outliers for Intervention=cal. Usually, researchers may consider to drop outliers since they have significant impact on model fitting, but since here outliers form a cluster approximately, removing outliers can be subtle here, it is recommended that outliers are retained.



## Formal analysis
This section outlines general strategies for quantifying and comparing the impacts of five different interventions on the consumption of zero-calorie and sugary beverages across three sites. 

It is recommended to implement Linear mixed model (LMM) to approach the problems. A LMM is a statistical model that can incorporate both fixed effects and random effects. In particular, the fixed effects are typically the primary variables of interests that are representative of the entire population. On the other hand, the random effects are parameters that vary across different hierarchical levels of the data, such as individual variability or other group effects. 

In this study, the primary focus is to analyze the effects of the interventions, and therefore it is suggested to put intervention as fixed effect. In addition, days of the week should also be treated as fixed effect because the weekly fluctuation applies to the all three sites. To allow for a more comprehensive understanding of the intervention effects, it is advised to also include sites and total dailiy beverage consumption as random effects in the model. These covariates control for variability that might influence the patterns of zero-calorie/sugary beverage consumption. For instance, site differences (which is the secondary statistical question) may impact purchasing behavior due to demographic variations between urban and suburban areas. Similarly, accounting for total daily beverage consumption helps isolate the effect of human traffic on zero-calorie/sugary beverage choices. This comprehensive approach ensures that the analysis accurately reflects the interventions' impact, controlling for external factors.

Hence, the proposed models are $ZeroCal \sim Intervention + DofW + Count + Site + Total$ and $Sugary \sim Intervention + DofW + Count + Site + Total$, where $Intervention$ and $DofW$ are treated as fixed effects, and the others as random effects. Specifically, the coefficients for $Intervention$ measures the effects of corresponding interventions to the response variables, while controlling the covariates in the model. Interpreting the coefficients for $Intervention$ is crucial for answering the statistical question.


## Conclusions



## References



## Statistical Appendix
